# Response to nitrogen addition reveals metabolic and ecological strategies of soil bacteria
#### R scripts 

#### Install and run R packages
```{r Load libraries and data}
library(phyloseq)  
library(ggplot2)
library(plyr)
library(scales)
library(reshape)
library(reshape2)
library(RColorBrewer)
library(grid)
library(dplyr)
library(vegan)
library(mpmcorrelogram)

uzdir <- "/file path/Samad_Response_to_N_Mol_Ecol_2017/"
otutable_biom_file <- paste(uzdir, "113_soils_merged_otu_table.json", sep = "")
map_file <- paste(uzdir, "113_cDNA_DNA_mapping file.txt", sep = "")
```

```{r Create phyloseq object}
# Now import the .biom-formatted otu_table-tax_table file.
biom_otu_tax <- import_biom(otutable_biom_file)

# Add sample data to the dataset using merge
bmsd <- import_qiime_sample_data(map_file)


##### Merge into phyloseq format
SS1_phyloseq <- merge_phyloseq(biom_otu_tax, bmsd)
SS1_phyloseq
sample_sums(SS1_phyloseq)

###### Save original phyloseq file
SS1_phyloseq0 = SS1_phyloseq
```

##### Create average result for multiple rarefaction by transforming data using (divide by 10) and check counts per sample
```{r Create average result for multiple rarefaction by transforming data using (divide by 10), results='markup'}
SS1_phyloseq = transform_sample_counts(SS1_phyloseq, function(x) x/10)
sample_sums(SS1_phyloseq)
```

##### Round and confirm count number
```{r Round and confirm count number, results='markup'}
SS1_phyloseq = transform_sample_counts(SS1_phyloseq, round)
sample_sums(SS1_phyloseq)
#check that all numbers are integers
otu_table_df<- as.data.frame(otu_table(SS1_phyloseq))
```

##### Check that all OTUs have representative counts  
```{r identify taxa with only zeros, results='markup', echo=TRUE}
any(taxa_sums(SS1_phyloseq) < 1)
sum(taxa_sums(SS1_phyloseq) < 1)
```

##### Prune taxa with less than 1 count and check taxa numbers again
```{r  Save original file and create new file with only present (no zeroes) taxa, results='markup', echo=TRUE}

#Create new file with only present (no zeroes) taxa

SS1_phyloseq = prune_taxa(taxa_sums(SS1_phyloseq) > 1, SS1_phyloseq)
any(sample_sums(SS1_phyloseq) < 1)
sum(taxa_sums(SS1_phyloseq) < 1)
```

##### Compare sequences per sample or OTU
```{r Compare sequences per sample or OTU}
readsumsdf = data.frame(nreads = sort(taxa_sums(SS1_phyloseq), TRUE), sorted = 1:ntaxa(SS1_phyloseq), 
                        type = "OTU")
readsumsdf = rbind(readsumsdf, data.frame(nreads = sort(sample_sums(SS1_phyloseq), 
                                                        TRUE), sorted = 1:nsamples(SS1_phyloseq), type = "Samples"))
title = "Total number of reads"
p = ggplot(readsumsdf, aes(x = sorted, y = nreads)) + geom_bar(stat = "identity")
p + ggtitle(title) + scale_y_log10() + facet_wrap(~type, 1, scales = "free")
```

```{r}
sample_variables(SS1_phyloseq)
```

```{r Attached OTU ID}
tax_table(SS1_phyloseq) <- cbind(tax_table(SS1_phyloseq), OTU=taxa_names(SS1_phyloseq))
```

```{r Rename Ranks}
colnames(tax_table(SS1_phyloseq)) = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species", "OTU" )
```


##### Subset samples by experiment
```{r}
cDNA_phyloseq = subset_samples(SS1_phyloseq, Type == "cDNA")

DNA_phyloseq = subset_samples(SS1_phyloseq, Type == "DNA")

#or excluding field samples
cDNA_phyloseq2 = subset_samples(SS1_phyloseq, Type2 == "cDNA")

DNA_phyloseq2 = subset_samples(SS1_phyloseq, Type2 == "DNA")

#Subset no Field sample 
DNA_cDNA_phyloseq = subset_samples(SS1_phyloseq, Type3 == "DNA_cDNA")

#Excluding Day 0
DNA_HM_phyloseq = subset_samples(DNA_phyloseq, Group == "HM")
DNA_LM_phyloseq = subset_samples(DNA_phyloseq, Group == "LM")
cDNA_HM_phyloseq = subset_samples(cDNA_phyloseq, Group == "HM")
cDNA_LM_phyloseq = subset_samples(cDNA_phyloseq, Group == "LM")

```


##### Plot alpha diversity
```{r, echo=FALSE}
#####NB. if you want to calculate and export diversity data then use the below commands
#####name_cDNA= estimate_richness(cDNA_phyloseq, split = TRUE, measures = NULL)
#####write.table(name_cDNA, "/file path/name_cDNA.txt", sep="\t")

#####name_DNA= estimate_richness(DNA_phyloseq, split = TRUE, measures = NULL)
#####write.table(name_DNA, "/file path/name_DNA.txt", sep="\t")

#### Plot alpha diversity at 16S rRNA level
##### richness plots
plot_richness(cDNA_phyloseq2, measures = c("Observed"), x = "Day") + geom_text(aes(label = " "), size = 4) + geom_boxplot(alpha = 1) + theme_bw() + theme(axis.text.x = element_text(hjust = 0.5, size = 16, color = "black"), axis.text.y = element_text(size = 16, color = "black"), axis.title.y = element_text(vjust = 1, size = 16),
        axis.title.x = element_text(size = 16), strip.text.x = element_text(size = 16)) + facet_grid(.~Treatment, drop = TRUE, space = "free", scales = "free") + ylab("Observed")

#### Shannon diversity
plot_richness(cDNA_phyloseq2, measures = c("Shannon"), x = "Day") + geom_text(aes(label = " "), size = 4) + geom_boxplot(alpha = 1) + theme_bw() + theme(axis.text.x = element_text(hjust = 0.5, size = 16, color = "black"), axis.text.y = element_text(size = 16, color = "black"), axis.title.y = element_text(vjust = 1, size = 16),
        axis.title.x = element_text(size = 16), strip.text.x = element_text(size = 16)) + facet_grid(.~Treatment, drop = TRUE, space = "free", scales = "free") + ylab("Shannon Diversity")

##### Plot alpha diversity at 16s rDNA level
#### richness plots
plot_richness(DNA_phyloseq, measures = c("Observed"), x = "Day") + geom_text(aes(label = " "), size = 4) + geom_boxplot(alpha = 1) + theme_bw() + theme(axis.text.x = element_text(hjust = 0.5, size = 16, color = "black"), axis.text.y = element_text(size = 16, color = "black"), axis.title.y = element_text(vjust = 1, size = 16),
        axis.title.x = element_text(size = 16), strip.text.x = element_text(size = 16)) + facet_grid(.~Treatment, drop = TRUE, space = "free", scales = "free") + ylab("Observed")

#### Shannon diversity
plot_richness(DNA_phyloseq, measures = c("Shannon"), x = "Day") + geom_text(aes(label = " "), size = 4) + geom_boxplot(alpha = 1) + theme_bw() + theme(axis.text.x = element_text(hjust = 0.5, size = 16, color = "black"), axis.text.y = element_text(size = 16, color = "black"), axis.title.y = element_text(vjust = 1, size = 16),
        axis.title.x = element_text(size = 16), strip.text.x = element_text(size = 16)) + facet_grid(.~Treatment, drop = TRUE, space = "free", scales = "free") + ylab("Shannon Diversity")
```

##### Subset samples by experiment (optional)
```{r}
Control_phyloseq = subset_samples(sainur_phyloseq, Treatment == "Control")

No_Urea_10kPa_phyloseq = subset_samples(sainur_phyloseq, Treatment == "No_Urea_10kPa")

No_Urea_1.0kPa_phyloseq = subset_samples(sainur_phyloseq, Treatment == "No_Urea_1.0kPa")

Urea_1.0kPa_phyloseq = subset_samples(sainur_phyloseq, Treatment == "Urea_1.0kPa")

Urea_10kPa_phyloseq = subset_samples(sainur_phyloseq, Treatment == "Urea_10kPa")

```

##### Figure S8 (Phylum level changes (relative abundance) in genome (16S rDNA) and transcript (16S rRNA) levels representing relative contribution >1% of all detected phyla (based on OTUs clustered at 97% sequence similarity)
```{r Figure S8}
sainur_phylum <- DNA_cDNA_phyloseq %>% tax_glom(taxrank = "Phylum") %>%   
# agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>%
# Transform to rel. abundance
  psmelt() %>%                
# Melt to long format
  filter(Abundance > 0.01) %>%                         
# Filter out low abundance taxa
  arrange(Phylum)  
  
phylum_colors <- c(
  "#CBD588", "#5F7FC7", "orange","#DA5724", "#508578", "#CD9BCD",
   "#AD6F3B", "#673770","#C84248", "#652926","#D14285", 
  "black", "#8A7C64", "#599861","#8569D5", "#5E738F")

color_labels=c("Acidobacteria", "Actinobacteria", "Bacteroidetes","Candidate division WS3","Chloroflexi","Cyanobacteria", "Firmicutes", "Gemmatimonadetes","Nitrospirae","Planctomycetes", "Proteobacteria", "Thaumarchaeota","Verrucomicrobia") 

#Day as factor
n_Day  <- factor(sainur_phylum$Day2)

library(Rmisc)
tg <- summarySE(sainur_phylum, measurevar="Abundance", groupvars=c("Sites2","n_Day","Phylum", "Treatment4"))
tg
require(grid)
require(plyr)
library(ggplot2)

pd <- position_dodge(0.001) # move them .05 to the left and right

f_DNA_cDNA=ggplot(tg, aes(x=n_Day, y=Abundance, group=Treatment4, fill = Phylum)) + 

  
  facet_wrap(~Treatment4, ncol = 2, scales = "free_y") +
  
 geom_bar(stat = "identity") +
  scale_fill_manual(values = phylum_colors,labels=color_labels)  +
  xlab("Day")+
  theme(axis.title.x = element_blank()) + 
   theme_bw()+
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.text.x = element_text(size=12, angle=0)) +
  guides(fill = guide_legend(reverse = T, keywidth = 1, keyheight = 1)) +
  ylab("Relative abundance (Phyla > 1%) \n") + 
 
  ggtitle("") 
f_DNA_cDNA


a= b+ scale_fill_manual(name="Phylum",values = c("grey26", "royalblue", "darkorange","chartreuse3",  "red", "#D3D93E", "#38333E", "#ff3cbb","cyan2", "darkgreen", "deepskyblue", "mediumorchid3","#89C5DA", "#DA5724", "#74D944", "#C84248", "#673770", "#D3D93E", "#38333E", "#508578", "#D7C1B1", "#689030", "#AD6F3B", "#CD9BCD", "#D14285", "#6DDE88", "#CD9BCD", "#D14285", "#6DDE88", "#652926", "#7FDCC0", "#8569D5", "#5E738F", "#D1A33D", "#8A7C64", "#599861", "blue4", "yellow1", "violetred", "#CD9BCD", "#D14285", "#6DDE88", "#652926", "#7FDCC0", "#8569D5", "#5E738F", "#D1A33D", "#8A7C64", "#599861", "blue4", "yellow1", "violetred", "#990000", "#99CC00", "#003300", "#00CCCC", "#9966CC", "#993366", "#990033", "#4863A0", "#000033", "#330000", "#00CC99", "#00FF33", "#00CCFF", "#FF9933", "#660066", "#FF0066", "slateblue4", "tan4", "tomato4", "steelblue", "springgreen4", "snow4", "slategray2", "plum1", "yellow", "sienna"),
                        
labels=c("Acidobacteria", "Actinobacteria", "Bacteroidetes","Chloroflexi", "Firmicutes", "Gemmatimonadetes","Planctomycetes", "Proteobacteria", "Thaumarchaeota","Verrucomicrobia") )

```

##### Figure 2b (16S rDNA & 16S rRNA)
``` {r Figure 2b}

title = "DNA"
#par(mfrow=c(2,1))
DNA.ord = ordinate( DNA_phyloseq, "NMDS", "bray")
p = plot_ordination(DNA_phyloseq, DNA.ord, color = "Day", shape = "Treatment")+
  scale_color_manual(values = c("#a65628", "red", "#ffae19",
    "#4daf4a", "#1919ff", "darkorchid3", "magenta"))
p = p + geom_point(size = 4, alpha = 0.7) + theme_bw()+theme(text = element_text(size = 17))+ggtitle(title)
#p


#Stress plot (Figure S4)
stressplot(DNA.ord)
DNA.ord


title = "cDNA"
RNA.ord = ordinate(cDNA_phyloseq, "NMDS", "bray")
q = plot_ordination(cDNA_phyloseq, RNA.ord, color = "Day", shape = "Treatment")+
  scale_color_manual(values = c("#a65628", "red", "#ffae19",
    "#4daf4a", "#1919ff", "darkorchid3", "magenta"))
q = q + geom_point(size = 4, alpha = 0.7) + theme_bw()+theme(text = element_text(size = 17))+ggtitle(title)

#q
#Stress plot (Figure S4)
stressplot(RNA.ord)
RNA.ord


library(grid)
multiplot(p, q,cols=1)
```

##### Adonis test
``` {r Adonis test }
library("phyloseq")
library ("vegan")
all_OTUs = names(sort(taxa_sums(DNA_phyloseq2), TRUE))
GPex = prune_taxa(all_OTUs, DNA_phyloseq2)
GPdist = phyloseq::distance(GPex, "bray")
GPNMDS = ordinate(GPex, "NMDS", GPdist)
adonis(GPdist ~ Treatment1, as(sample_data(GPex), "data.frame"))

#####results:

#           Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
#Treatment1  1    2.1149  2.1149  18.045 0.26135  0.001 ***
#Residuals  51    5.9773  0.1172         0.73865           
#Total      52    8.0922                 1.00000  



##cDNA (16S rRNA)
library("phyloseq")
library ("vegan")
all_OTUs = names(sort(taxa_sums(cDNA_phyloseq2), TRUE))
SSex = prune_taxa(all_OTUs, cDNA_phyloseq2)
SSdist = phyloseq::distance(SSex, "bray")
SSNMDS = ordinate(SSex, "NMDS", SSdist)
adonis(SSdist ~ Treatment1, as(sample_data(SSex), "data.frame"))

#######results:
#           Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
#Treatment1  1    4.0404  4.0404   26.27 0.33563  0.001 ***
#Residuals  52    7.9978  0.1538         0.66437           
#Total      53   12.0381                 1.00000    

````

##### Mantel Correlation
```{r Figure 2c and S6 }
###subset samples by Treatments (DNA)

No_Urea_10kPa_phyloseq = subset_samples(DNA_phyloseq , Treatment_R1 == "No_Urea_10kPa")

No_Urea_1.0kPa_phyloseq = subset_samples(DNA_phyloseq , Treatment_R1 == "No_Urea_1.0kPa")

Urea_1.0kPa_phyloseq = subset_samples(DNA_phyloseq, Treatment_R1 == "Urea_1.0kPa")

Urea_10kPa_phyloseq = subset_samples(DNA_phyloseq, Treatment_R1 == "Urea_10kPa")

###subset samples by Treatments (cDNA)

No_Urea_10kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq , Treatment_R1 == "No_Urea_10kPa")

No_Urea_1.0kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq , Treatment_R1 == "No_Urea_1.0kPa")

Urea_1.0kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq, Treatment_R1 == "Urea_1.0kPa")

Urea_10kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq, Treatment_R1 == "Urea_10kPa")

## DNA: HM+Urea
library(vegan)
library(mpmcorrelogram)
#x.dis:Multivariate distance (or similarity) matrices
Bry_urea_HM <- as.matrix(phyloseq::distance(Urea_1.0kPa_phyloseq, method="bray"))
#GeoDis or timeDis
df_time <- as(sample_data(Urea_1.0kPa_phyloseq ), "data.frame")
Day.dist_HM <- dist(df_time$Day_order)
dist.matrix_HM <- as.matrix(Day.dist_HM)

par(mfrow=c(2,2))


x=mpmcorrelogram(Bry_urea_HM,Day.dist_HM, zdis=NULL, method = "spearman",
alfa = 0.05, nclass =NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(x, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=-0.5,"HM+N",cex=1,
        col="black",bty = "n")
legend(x=0,y=0.5,"16SrDNA",cex=1, col="black",bty = "n")

### LM+Urea
Bry_urea_LM <- as.matrix(phyloseq::distance(Urea_10kPa_phyloseq , method="bray"))
df_time <- as(sample_data(Urea_10kPa_phyloseq ), "data.frame")
Day.dist_LM <- dist(df_time$Day_order)
dist.matrix_LM <- as.matrix(Day.dist_LM)

y=mpmcorrelogram(Bry_urea_LM,dist.matrix_LM, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(y, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=-0.5,"LM+N",cex=1,
        col="black",bty = "n")
        legend(x=0,y=0.5,"16rDNA",cex=1,col="black",bty = "n")

#cDNA
Bry_urea_HM_cDNA <- as.matrix(phyloseq::distance(Urea_1.0kPa_phyloseq_cDNA , method="bray"))
df_time <- as(sample_data(Urea_1.0kPa_phyloseq_cDNA ), "data.frame")
Day.dist_HM_cDNA <- dist(df_time$Day_order)
dist.matrix_HM <- as.matrix(Day.dist_HM_cDNA)

z=mpmcorrelogram(Bry_urea_HM_cDNA,Day.dist_HM_cDNA, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(z, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=-0.5,"HM+N",cex=1,
        col="black",bty = "n")
    legend(x=0,y=0.5,"16rRNA",cex=1,col="black",bty = "n")

### LM+Urea
Bry_urea_LM_cDNA <- as.matrix(phyloseq::distance(Urea_10kPa_phyloseq_cDNA , method="bray"))
df_time <- as(sample_data(Urea_10kPa_phyloseq_cDNA ), "data.frame")
Day.dist_LM_cDNA <- dist(df_time$Day_order)
dist.matrix_LM <- as.matrix(Day.dist_HM_cDNA)

z=mpmcorrelogram(Bry_urea_LM_cDNA,Day.dist_LM_cDNA, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(z, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=-0.5,"LM+N",cex=1,
        col="black",bty = "n")

 legend(x=0,y=0.5,"16rRNA",cex=1,col="black",bty = "n")

#NB.Similar results were observed for second (R2) and third replicates (R3)
 
####Second replicate R2

No_Urea_10kPa_phyloseq = subset_samples(DNA_phyloseq , Treatment_R2 == "No_Urea_10kPa")

No_Urea_1.0kPa_phyloseq = subset_samples(DNA_phyloseq , Treatment_R2 == "No_Urea_1.0kPa")

Urea_1.0kPa_phyloseq = subset_samples(DNA_phyloseq, Treatment_R2 == "Urea_1.0kPa")

Urea_10kPa_phyloseq = subset_samples(DNA_phyloseq, Treatment_R2 == "Urea_10kPa")

###subset samples by Treatments (cDNA)

No_Urea_10kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq , Treatment_R2 == "No_Urea_10kPa")

No_Urea_1.0kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq , Treatment_R2 == "No_Urea_1.0kPa")

Urea_1.0kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq, Treatment_R2 == "Urea_1.0kPa")

Urea_10kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq, Treatment_R2 == "Urea_10kPa")

## DNA: HM+Urea
library(vegan)
library(mpmcorrelogram)
#x.dis:Multivariate distance (or similarity) matrices
Bry_urea_HM <- as.matrix(phyloseq::distance(Urea_1.0kPa_phyloseq, method="bray"))
#GeoDis or timeDis
df_time <- as(sample_data(Urea_1.0kPa_phyloseq ), "data.frame")
Day.dist_HM <- dist(df_time$Day_order)
dist.matrix_HM <- as.matrix(Day.dist_HM)

par(mfrow=c(2,2))


x=mpmcorrelogram(Bry_urea_HM,Day.dist_HM, zdis=NULL, method = "spearman",
alfa = 0.05, nclass =NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(x, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=0.9,"HM+N",cex=1,
        col="black",bty = "n")
legend(x=0,y=0.5,"16rDNA",cex=1, col="black",bty = "n")

### LM+Urea
Bry_urea_LM <- as.matrix(phyloseq::distance(Urea_10kPa_phyloseq , method="bray"))
df_time <- as(sample_data(Urea_10kPa_phyloseq ), "data.frame")
Day.dist_LM <- dist(df_time$Day_order)
dist.matrix_LM <- as.matrix(Day.dist_LM)

y=mpmcorrelogram(Bry_urea_LM,dist.matrix_LM, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(y, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=0.9,"LM+N",cex=1,
        col="black",bty = "n")
        legend(x=0,y=0.5,"16rDNA",cex=1,col="black",bty = "n")

#cDNA
Bry_urea_HM_cDNA <- as.matrix(phyloseq::distance(Urea_1.0kPa_phyloseq_cDNA , method="bray"))
df_time <- as(sample_data(Urea_1.0kPa_phyloseq_cDNA ), "data.frame")
Day.dist_HM_cDNA <- dist(df_time$Day_order)
dist.matrix_HM <- as.matrix(Day.dist_HM_cDNA)

z=mpmcorrelogram(Bry_urea_HM_cDNA,Day.dist_HM_cDNA, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(z, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=0.9,"HM+N",cex=1,
        col="black",bty = "n")
    legend(x=0,y=0.5,"16rRNA",cex=1,col="black",bty = "n")

### LM+Urea
Bry_urea_LM_cDNA <- as.matrix(phyloseq::distance(Urea_10kPa_phyloseq_cDNA , method="bray"))
df_time <- as(sample_data(Urea_10kPa_phyloseq_cDNA ), "data.frame")
Day.dist_LM_cDNA <- dist(df_time$Day_order)
dist.matrix_LM <- as.matrix(Day.dist_HM_cDNA)

z=mpmcorrelogram(Bry_urea_LM_cDNA,Day.dist_LM_cDNA, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(z, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "rM", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=0.9,"LM+N",cex=1,
        col="black",bty = "n")
  legend(x=0,y=0.5,"16rRNA",cex=1,col="black",bty = "n")



###Third replicate R3 

No_Urea_10kPa_phyloseq = subset_samples(DNA_phyloseq , Treatment_R3 == "No_Urea_10kPa")

No_Urea_1.0kPa_phyloseq = subset_samples(DNA_phyloseq , Treatment_R3 == "No_Urea_1.0kPa")

Urea_1.0kPa_phyloseq = subset_samples(DNA_phyloseq, Treatment_R3 == "Urea_1.0kPa")

Urea_10kPa_phyloseq = subset_samples(DNA_phyloseq, Treatment_R3 == "Urea_10kPa")
###subset samples by Treatments (cDNA)

No_Urea_10kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq , Treatment_R3 == "No_Urea_10kPa")

No_Urea_1.0kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq , Treatment_R3 == "No_Urea_1.0kPa")

Urea_1.0kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq, Treatment_R3 == "Urea_1.0kPa")

Urea_10kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq, Treatment_R3 == "Urea_10kPa")

## DNA: HM+Urea
library(vegan)
library(mpmcorrelogram)
#x.dis:Multivariate distance (or similarity) matrices
Bry_urea_HM <- as.matrix(phyloseq::distance(Urea_1.0kPa_phyloseq, method="bray"))
#GeoDis or timeDis
df_time <- as(sample_data(Urea_1.0kPa_phyloseq ), "data.frame")
Day.dist_HM <- dist(df_time$Day_order)
dist.matrix_HM <- as.matrix(Day.dist_HM)

par(mfrow=c(2,2))



x=mpmcorrelogram(Bry_urea_HM,Day.dist_HM, zdis=NULL, method = "spearman",
alfa = 0.05, nclass =NULL, breaks = c(1,2,3,4,5,6),
permutations = 9999, simil = T,
plot = F, print = TRUE)
plot(x, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=0.9,"HM+N",cex=1,
        col="black",bty = "n")
legend(x=0,y=0.5,"16rDNA",cex=1, col="black",bty = "n")

### LM+Urea
Bry_urea_LM <- as.matrix(phyloseq::distance(Urea_10kPa_phyloseq , method="bray"))
df_time <- as(sample_data(Urea_10kPa_phyloseq ), "data.frame")
Day.dist_LM <- dist(df_time$Day_order)
dist.matrix_LM <- as.matrix(Day.dist_LM)

y=mpmcorrelogram(Bry_urea_LM,dist.matrix_LM, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(y, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=0.9,"LM+N",cex=1,
        col="black",bty = "n")
        legend(x=0,y=0.5,"16rDNA",cex=1,col="black",bty = "n")

#cDNA
Bry_urea_HM_cDNA <- as.matrix(phyloseq::distance(Urea_1.0kPa_phyloseq_cDNA , method="bray"))
df_time <- as(sample_data(Urea_1.0kPa_phyloseq_cDNA ), "data.frame")
Day.dist_HM_cDNA <- dist(df_time$Day_order)
dist.matrix_HM <- as.matrix(Day.dist_HM_cDNA)

z=mpmcorrelogram(Bry_urea_HM_cDNA,Day.dist_HM_cDNA, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(z, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=0.9,"HM+N",cex=1,
        col="black",bty = "n")
    legend(x=0,y=0.5,"16rRNA",cex=1,col="black",bty = "n")

### LM+Urea
Bry_urea_LM_cDNA <- as.matrix(phyloseq::distance(Urea_10kPa_phyloseq_cDNA , method="bray"))
df_time <- as(sample_data(Urea_10kPa_phyloseq_cDNA ), "data.frame")
Day.dist_LM_cDNA <- dist(df_time$Day_order)
dist.matrix_LM <- as.matrix(Day.dist_HM_cDNA)

z=mpmcorrelogram(Bry_urea_LM_cDNA,Day.dist_LM_cDNA, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4,5,6),
permutations = 999, simil = T,
plot = F, print = TRUE)
plot(z, pch = c(19, 1), xlim = c(0.5,5.5), ylim = c(-1,1),
ylab = "rM", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=4,y=0.9,"LM+N",cex=1,
        col="black",bty = "n")
legend(x=0,y=0.5,"16rRNA",cex=1,col="black",bty = "n")


###NO urea treatment (replicate 1, Figure S6)
# subset samples 
No_Urea_10kPa_phyloseq = subset_samples(DNA_phyloseq , Treatment_R1 == "No_Urea_10kPa")

No_Urea_1.0kPa_phyloseq = subset_samples(DNA_phyloseq , Treatment_R1 == "No_Urea_1.0kPa")
No_Urea_10kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq , Treatment_R1 == "No_Urea_10kPa")

No_Urea_1.0kPa_phyloseq_cDNA = subset_samples(cDNA_phyloseq , Treatment_R1 == "No_Urea_1.0kPa")

## DNA: HM+Urea
library(vegan)
library(mpmcorrelogram)
#x.dis:Multivariate distance (or similarity) matrices
Bry_urea_HM <- as.matrix(phyloseq::distance(No_Urea_1.0kPa_phyloseq, method="bray"))
#GeoDis or timeDis
df_time <- as(sample_data(No_Urea_1.0kPa_phyloseq), "data.frame")
Day.dist_HM <- dist(df_time$Day_order)
dist.matrix_HM <- as.matrix(Day.dist_HM)

par(mfrow=c(2,2))


x=mpmcorrelogram(Bry_urea_HM,Day.dist_HM, zdis=NULL, method = "spearman",
alfa = 0.05, nclass =NULL, breaks = c(1,2,3,4),
permutations = 999, simil = F,
plot = F, print = TRUE)
plot(x, pch = c(19, 1), xlim = c(0.5,4.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=3,y=-0.5,"HM-N",cex=1,
        col="black",bty = "n")
legend(x=0,y=-0.5,"16rDNA",cex=1, col="black",bty = "n")

### LM+Urea
Bry_urea_LM <- as.matrix(phyloseq::distance(No_Urea_10kPa_phyloseq, method="bray"))
df_time <- as(sample_data(No_Urea_10kPa_phyloseq ), "data.frame")
Day.dist_LM <- dist(df_time$Day_order)
dist.matrix_LM <- as.matrix(Day.dist_LM)

y=mpmcorrelogram(Bry_urea_LM,dist.matrix_LM, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4),
permutations = 999, simil = F,
plot = F, print = TRUE)
plot(y, pch = c(19, 1), xlim = c(0.5,4.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=3,y=-0.5,"LM-N",cex=1,
        col="black",bty = "n")
        legend(x=0,y=-0.5,"16rDNA",cex=1,col="black",bty = "n")

#cDNA
Bry_urea_HM_cDNA <- as.matrix(phyloseq::distance(No_Urea_1.0kPa_phyloseq_cDNA , method="bray"))
df_time <- as(sample_data(No_Urea_1.0kPa_phyloseq_cDNA ), "data.frame")
Day.dist_HM_cDNA <- dist(df_time$Day_order)
dist.matrix_HM <- as.matrix(Day.dist_HM_cDNA)

z=mpmcorrelogram(Bry_urea_HM_cDNA,Day.dist_HM_cDNA, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4),
permutations = 999, simil = F,
plot = F, print = TRUE)
plot(z, pch = c(19, 1), xlim = c(0.5,4.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=3,y=-0.5,"HM-N",cex=1,
        col="black",bty = "n")
    legend(x=0,y=-0.5,"16rRNA",cex=1,col="black",bty = "n")

### LM+Urea
Bry_urea_LM_cDNA <- as.matrix(phyloseq::distance(No_Urea_10kPa_phyloseq_cDNA , method="bray"))
df_time <- as(sample_data(No_Urea_10kPa_phyloseq_cDNA ), "data.frame")
Day.dist_LM_cDNA <- dist(df_time$Day_order)
dist.matrix_LM <- as.matrix(Day.dist_HM_cDNA)

z=mpmcorrelogram(Bry_urea_LM_cDNA,Day.dist_LM_cDNA, zdis=NULL, method = "spearman",
alfa = 0.05, nclass = NULL, breaks = c(1,2,3,4),
permutations = 999, simil = F,
plot = F, print = TRUE)
plot(z, pch = c(19, 1), xlim = c(0.5,4.5), ylim = c(-1,1),
ylab = "Mantel correlation coefficient", xlab = "Distance class (Time series)", alfa = 0.05)
abline(h=0,col="gray", lty =2)
legend(x=3,y=-0.5,"LM-N",cex=1,
        col="black",bty = "n")
  legend(x=0,y=-0.5,"16rRNA",cex=1,col="black",bty = "n")

```


```{r SIMPER analysis}
# SIMPER analysis (Groups LM and HM, Treatment= = +N and -N)
#Create new file with only present (no zeroes) taxa
DNA_HM_phyloseq = prune_taxa(taxa_sums(DNA_HM_phyloseq) > 1, DNA_HM_phyloseq)
any(sample_sums(DNA_HM_phyloseq) < 1)
sum(taxa_sums(DNA_HM_phyloseq) < 1)

#1. SIMPER (HM_DNA)
phyloseq_mat <- data.frame(otu_table(DNA_HM_phyloseq))
phyloseq_matt <- t(phyloseq_mat)
phyloseq_sd <- data.frame(sample_data(DNA_HM_phyloseq))
(sim_HM_DNA <- with(phyloseq_sd, simper(phyloseq_matt, Treatment1),permutations = 999))
summary(sim_HM_DNA)
result_HM <- summary(sim_HM_DNA) 
lapply(sim_HM_DNA, FUN = function(x){x$overall})


#export data
sim_HM_DNA <- result_HM$"Urea_No Urea"
write.table(sim_HM_DNA, "/file path/HM_DNA.txt",sep="\t")

#2. SIMPER (LM_DNA)
DNA_LM_phyloseq = prune_taxa(taxa_sums(DNA_LM_phyloseq) > 1, DNA_LM_phyloseq)
any(sample_sums(DNA_LM_phyloseq) < 1)
sum(taxa_sums(DNA_LM_phyloseq) < 1)

#SIMPER (LM_DNA)
phyloseq_mat <- data.frame(otu_table(DNA_LM_phyloseq))
phyloseq_matt <- t(phyloseq_mat)
phyloseq_sd <- data.frame(sample_data(DNA_LM_phyloseq))
(sim_LM_DNA <- with(phyloseq_sd, simper(phyloseq_matt, Treatment1),permutations = 999))
summary(sim_LM_DNA)
result_LH <- summary(sim_LM_DNA)
lapply(sim_LM_DNA, FUN = function(x){x$overall})

#Export file
sim_LM_DNA <- result_LH $"No Urea_Urea"
write.table(sim_LM_DNA, "/file path/LM_DNA.txt",sep="\t")


#cDNA
cDNA_HM_phyloseq = prune_taxa(taxa_sums(cDNA_HM_phyloseq) > 1, cDNA_HM_phyloseq)
any(sample_sums(cDNA_HM_phyloseq) < 1)
sum(taxa_sums(cDNA_HM_phyloseq) < 1)

#SIMPER (HM_cDNA)
phyloseq_mat <- data.frame(otu_table(cDNA_HM_phyloseq))
phyloseq_matt <- t(phyloseq_mat)
phyloseq_sd <- data.frame(sample_data(cDNA_HM_phyloseq))
(sim_HM_cDNA <- with(phyloseq_sd, simper(phyloseq_matt, Treatment1),permutations = 999))
summary(sim_HM_cDNA)
result_HM_ccDNA <- summary(sim_HM_cDNA) 
lapply(sim_HM_cDNA, FUN = function(x){x$overall})

#export data
sim_HM_cDNA <- result_HM_ccDNA$"No Urea_Urea"
write.table(sim_HM_DNA, "/file path/HM_cDNA.txt",sep="\t")

#2. SIMPER (LM_cDNA)
cDNA_LM_phyloseq = prune_taxa(taxa_sums(cDNA_LM_phyloseq) > 1, cDNA_LM_phyloseq)
any(sample_sums(cDNA_LM_phyloseq) < 1)
sum(taxa_sums(cDNA_LM_phyloseq) < 1)

#SIMPER (LM_DNA)
phyloseq_mat <- data.frame(otu_table(cDNA_LM_phyloseq))
phyloseq_matt <- t(phyloseq_mat)
phyloseq_sd <- data.frame(sample_data(cDNA_LM_phyloseq))
(sim_LM_cDNA <- with(phyloseq_sd, simper(phyloseq_matt, Treatment1),permutations = 999))
summary(sim_LM_cDNA)
result_LH_cDNA <- summary(sim_LM_cDNA) 
lapply(sim_LM_cDNA, FUN = function(x){x$overall})


#Export
sim_LM_cDNA <- result_LH_cDNA $"Urea_No Urea"
write.table(sim_LM_DNA, "/Users/Sainur/Desktop/LM_cDNA.txt",sep="\t")

```

#### Pvclust tree using bray distance and including p values for each node. [AU (approximately unbiased) BP (bootstrap probability)]. Boxes mark clusters with 95% confidence
```{r Cluster tree with box marking 95percent confidence, Figure S5}
library(pvclust)
library("devtools") # used to source functions from the internet
source_url("https://raw.githubusercontent.com/nielshanson/mp_tutorial/master/taxonomic_analysis/code/pvclust_bcdist.R")
# alternatively, you might want to download the above function and load using source()
# e.g. source('/where/I/downloaded/pvclust_bcdist.R')
#extract otu information
otus <- otu_table(DNA_phyloseq)
otus_ss <- otus@.Data
fit <- pvclust(otus_ss, method.hclust="ward", method.dist="bray–curtis", n=1000)
   plot(fit) # dendogram with p values
pvrect(fit, alpha=.95)
```

```{r t-test}
#### t-test
##### independent 2-group t-test, #t.test(y1,y2) # where y1 and y2 are numeric
t_test<- read.csv("/file path/HM_rrn_t_test.csv", header=T)
t_test
t.test(t_test$Max.abundance_HM_H,t_test$Max.abundance_HM_L)
#(p=t = 3.2014, df = 30.656, p-value = 0.003178)
t.test(t_test$Fold.changes_HM_H,t_test$Fold.changes_HM_L)
#(t = 2.7129, df = 22.423, p-value = 0.01258)
t.test(t_test$Growth.rate_HM_H,t_test$Growth.rate_HM_L)
#(t = 0.58204, df = 23.923, p-value = 0.566)

t_test<- read.csv("/file path/LM_rrn_t_test.csv", header=T)

t_test
t.test(t_test$Max.abundance_LM_H,t_test$Max.abundance_LM_L)
#(t = 3.5551, df = 21.617, p-value = 0.00181)
t.test(t_test$Fold.changes_LM_H,t_test$Fold.changes_LM_L)
#(t = 3.1705, df = 14.42, p-value = 0.0066)
t.test(t_test$Growth.rate_LM_H,t_test$Growth.rate_LM_L)
#(t = 0.97296, df = 20.382, p-value = 0.342)


#best(HM/LM)_t
t_test<- read.csv("~/Documents/R_files/best_rrn_t_test.csv", header=T)

t.test(t_test$Growth_best_H,t_test$Growth_best_L)
#(t = 0.72294, df = 25.938, p-value = 0.4762)
t.test(t_test$Max_abundance_best_H,t_test$Max_abundance_best_L)
#(t = 3.7407, df = 32.934, p-value = 0.0006993)
t.test(t_test$Fold_changes_H,t_test$Fold_changes_L)
#(t = 3.5184, df = 22.455, p-value = 0.001893)
```

##### Michaelis_Menten Model (Figure 7)
```{r }
library(drc)
library(grid)
library(ggplot2)

mn_test<- read.csv("/file path/HM_mn_test.csv", header=T)
mn_test

model.drm <- drm (Growth_Day_HM ~ rrn_copy_HM, data = mn_test, fct = MM.3())

mml <- data.frame(rrn_copy_HM = seq(0, max(mn_test$rrn_copy_HM), length.out = 100))
mml$Growth_Day_HM <- predict(model.drm, newdata = mml)

#drawing graph in ggplot
Phylum = mn_test$Phylum
HM1=ggplot(mn_test, aes(x=rrn_copy_HM, y=Growth_Day_HM, color= Phylum)) +
  xlab("rrn copy no.") +
  ylab("Growth rate (per day)") +
  ggtitle("HM") +
  geom_point(alpha = 0.7, size =3) + 
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  geom_line(data = mml, aes(x = rrn_copy_HM, y = Growth_Day_HM), colour = "gray")
HM1
HM1= HM1 +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
HM1



###Maximum
model.drm <- drm (Max_abundance_HM ~ rrn_copy_HM, data = mn_test, fct = MM.2())

mml <- data.frame(rrn_copy_HM = seq(0, max(mn_test$rrn_copy_HM), length.out = 100))
mml$Max_abundance_HM <- predict(model.drm, newdata = mml)

#drawing graph in ggplot

Phylum = mn_test$Phylum
HM2= ggplot(mn_test, aes(x = rrn_copy_HM, y = Max_abundance_HM, color =Phylum)) +
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  xlab("rrn copy no.") +
  ylab("Max. abundance") +
  ggtitle("HM") +
  geom_point(alpha = 0.7, size = 3) +
  
 
geom_line(data = mml, aes(x = rrn_copy_HM, y = Max_abundance_HM), colour = "gray")


HM2= HM2 +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
HM2




#Foldchnages_HM
model.drm <- drm (Fold_changes_HM ~ rrn_copy_HM, data = mn_test, fct = MM.2())

mml <- data.frame(rrn_copy_HM = seq(0, max(mn_test$rrn_copy_HM), length.out = 100))
mml$Fold_changes_HM <- predict(model.drm, newdata = mml)

#drawing graph in ggplot

HM3= ggplot(mn_test, aes(x = rrn_copy_HM, y = Fold_changes_HM, color =Phylum)) +
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  xlab("rrn copy no.") +
  ylab("Fold change") +
  ggtitle("HM") +
  geom_point(alpha = 0.7, size=3) +
  geom_line(data = mml, aes(x = rrn_copy_HM, y = Fold_changes_HM), colour = "gray")
HM3

HM3= HM3 +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
HM3


###LM
LM_test<- read.csv("/file path/LM_mn_test .csv", header=T)
LM_test

model.drm <- drm (Growth_Day_LM ~ rrn_copy_LM, data = LM_test, fct = MM.2())

mml <- data.frame(rrn_copy_LM = seq(0, max(LM_test$rrn_copy_LM), length.out = 100))
mml$Growth_Day_LM <- predict(model.drm, newdata = mml)

#drawing graph in ggplot

LM1=ggplot(LM_test, aes(x = rrn_copy_LM, y = Growth_Day_LM, color=Phylum)) +
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  xlab("rrn copy no.") +
  ylab("Growth rate (per day)") +
  ggtitle("LM") +
  geom_point(alpha = 0.7, size= 3) +
  geom_line(data = mml, aes(x = rrn_copy_LM, y = Growth_Day_LM), colour = "gray")
LM1
LM1= LM1 +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
LM1


###Maximum
model.drm <- drm (Max_abundance_LM ~ rrn_copy_LM, data = LM_test, fct = MM.2())

mml <- data.frame(rrn_copy_LM = seq(0, max(LM_test$rrn_copy_LM), length.out = 100))
mml$Max_abundance_LM <- predict(model.drm, newdata = mml)

#drawing graph in ggplot

LM2= ggplot(LM_test, aes(x = rrn_copy_LM, y = Max_abundance_LM, color=Phylum)) +
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  xlab("rrn copy no.") +
  ylab("Max. abundance") +
  ggtitle("LM") +
  geom_point(alpha = 0.7, size= 3) +
  geom_line(data = mml, aes(x = rrn_copy_LM, y = Max_abundance_LM), colour = "gray")
LM2
LM2= LM2 +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
LM2

#Foldchnages_HM
model.drm <- drm (Fold_changes_LM ~ rrn_copy_LM, data = LM_test, fct = MM.2())

mml <- data.frame(rrn_copy_LM = seq(0, max(LM_test$rrn_copy_LM), length.out = 100))
mml$Fold_changes_LM <- predict(model.drm, newdata = mml)

#drawing graph in ggplot
LM3= ggplot(LM_test, aes(x = rrn_copy_LM, y = Fold_changes_LM, color = Phylum)) +
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  xlab("rrn copy no.") +
  ylab("Fold change") +
  ggtitle("LM") +
  geom_point(alpha = 0.7, size=3) +
  geom_line(data = mml, aes(x = rrn_copy_LM, y = Fold_changes_LM), colour = "gray")
LM3
LM3= LM3 +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
LM3


###BestLM/HM
best_test<- read.csv("/file path/best_LM_HM.csv", header=T)
best_test

model.drm <- drm (Growth_Day_best ~ rrn_copy_best, data = best_test, fct = MM.2())

mml <- data.frame(rrn_copy_best = seq(0, max(best_test$rrn_copy_best), length.out = 100))
mml$Growth_Day_best <- predict(model.drm, newdata = mml)

#drawing graph in ggplot
best1=ggplot(best_test, aes(x = rrn_copy_best, y = Growth_Day_best, color=Phylum)) +
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  xlab("rrn copy no.") +
  ylab("Growth rate (per day)") +
  ggtitle("Best (HM/LM)") +
  geom_point(alpha = 0.7, size= 3) +
  geom_line(data = mml, aes(x = rrn_copy_best, y = Growth_Day_best), colour = "gray")
best1
best1= best1 +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
best1

###Maximum
model.drm <- drm (Max_abundance_best ~ rrn_copy_best, data = best_test, fct = MM.2())

mml <- data.frame(rrn_copy_best = seq(0, max(best_test$rrn_copy_best), length.out = 100))
mml$Max_abundance_best <- predict(model.drm, newdata = mml)

#drawing graph in ggplot
best2= ggplot(best_test, aes(x = rrn_copy_best, y = Max_abundance_best, color=Phylum)) +
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  xlab("rrn copy no.") +
  ylab("Max. abundance") +
  ggtitle("Best (HM/LM)") +
  geom_point(alpha = 0.7, size= 3) +
  geom_line(data = mml, aes(x = rrn_copy_best, y = Max_abundance_best), colour = "gray")
best2
best2= best2 +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
best2

#Foldchnages_HM
model.drm <- drm (Fold_changes_best ~ rrn_copy_best, data = best_test, fct = MM.2())

mml <- data.frame(rrn_copy_best = seq(0, max(best_test$rrn_copy_best), length.out = 100))
mml$Fold_changes_best <- predict(model.drm, newdata = mml)

#drawing graph in ggplot
best3= ggplot(best_test, aes(x = rrn_copy_best, y = Fold_changes_best, color = Phylum)) +
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  xlab("rrn copy no.") +
  ylab("Fold change") +
  ggtitle("Best (HM/LM)") +
  geom_point(alpha = 0.7, size=3) +
  geom_line(data = mml, aes(x = rrn_copy_best, y = Fold_changes_best), colour = "gray")
best3
best3= best3 +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
best3


p1<- multiplot(HM1,HM2,HM3,LM1,LM2,LM3,best1,best2,best3, cols=3)

##NB: Multiplot function can loaded from here: http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/ 

```
#### Linear fit rrn (Figure S11)
```{r }
library(ggplot2)

mn_test<- read.csv("/file path/Linear model data/HM_linear_model.csv", header=T)
mn_test

mn.lm = lm(Growth_Day_HM~ log2.rrn , data=mn_test)
summary(mn.lm) 

#drawing graph in ggplot
Phylum = mn_test$Phylum
HM1=ggplot(mn_test, aes(x=log2.rrn, y=Growth_Day_HM)) +
  geom_smooth(method = "lm", color="black")+
  xlab("rrn copy no. (log2)") +
  ylab("Growth rate (per day)") +
  ggtitle("HM") +
  geom_point(aes(color = Phylum),alpha = 0.7, size =3) + 
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
 
HM1
 
###Maximum
max.lm = lm(Max_abundance_HM~ log2.rrn , data=mn_test)
summary(max.lm) 

#drawing graph in ggplot

Phylum = mn_test$Phylum
HM2=ggplot(mn_test, aes(x=log2.rrn, y=Max_abundance_HM)) +
  geom_smooth(method = "lm", color="black")+
  xlab("rrn copy no. (log2)") +
  ylab("Max. abundance ") +
  ggtitle("HM") +
  geom_point(aes(color = Phylum),alpha = 0.7, size =3) + 
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

HM2

#Foldchnages_HM

max.lm = lm(Fold_changes_HM~ log2.rrn , data=mn_test)
summary(max.lm) 

#drawing graph in ggplot

Phylum = mn_test$Phylum
HM3=ggplot(mn_test, aes(x=log2.rrn, y=Fold_changes_HM)) +
  geom_smooth(method = "lm", color="black")+
  xlab("rrn copy no. (log2)") +
  ylab("Fold change (abundance)") +
  ggtitle("HM") +
  geom_point(aes(color = Phylum),alpha = 0.7, size =3) + 
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

HM3

###LM
LM_test<- read.csv("/file path/LM_linear_model.csv", header=T)
LM_test

max.lm = lm(Growth_Day_LM~ log2rrn , data=LM_test)
summary(max.lm) 

#drawing graph in ggplot
Phylum = LM_test$Phylum
LM1=ggplot(LM_test, aes(x=log2rrn, y=Growth_Day_LM)) +
  geom_smooth(method = "lm", color="black")+
  xlab("rrn copy no. (log2)") +
  ylab("Growth rate (per day)") +
  ggtitle("LM") +
  geom_point(aes(color = Phylum),alpha = 0.7, size =3) + 
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

LM1

###Maximum
max.lm = lm(Max_abundance_LM~ log2rrn , data=LM_test)
summary(max.lm) 

#drawing graph in ggplot

Phylum = LM_test$Phylum
LM2=ggplot(LM_test, aes(x=log2rrn, y=Max_abundance_LM)) +
  geom_smooth(method = "lm", color="black")+
  xlab("rrn copy no. (log2)") +
  ylab("Max. abundance ") +
  ggtitle("LM") +
  geom_point(aes(color = Phylum),alpha = 0.7, size =3) + 
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

LM2


#Foldchnages_HM
max.lm = lm(Fold_changes_LM~ log2rrn , data=LM_test)
summary(max.lm) 

#drawing graph in ggplot
Phylum = LM_test$Phylum
LM3=ggplot(LM_test, aes(x=log2rrn, y=Fold_changes_LM)) +
  geom_smooth(method = "lm", color="black")+
  xlab("rrn copy no. (log2)") +
  ylab("Fold change (abundance)") +
  ggtitle("LM") +
  geom_point(aes(color = Phylum),alpha = 0.7, size =3) + 
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

LM3

###Best (LM/HM)
Best_test<- read.csv("/file path/best_linear_LM_HM.csv", header=T)
Best_test

bestG.lm = lm(Growth_Day_best~ log2rrn , data=Best_test)
summary(bestG.lm) 

#drawing graph in ggplot
Phylum = Best_test$Phylum
best1=ggplot(Best_test, aes(x=log2rrn, y=Growth_Day_best)) +
  geom_smooth(method = "lm", color="black")+
  xlab("rrn copy no. (log2)") +
  ylab("Growth rate (per day)") +
  ggtitle("Best (HM/LM)") +
  geom_point(aes(color = Phylum),alpha = 0.7, size =3) + 
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

best1

###Maximum
best_max.lm = lm(Max_abundance_best~ log2rrn , data=Best_test)
summary(best_max.lm) 

#drawing graph in ggplot

Phylum = Best_test$Phylum
best2=ggplot(Best_test, aes(x=log2rrn, y=Max_abundance_best)) +
  geom_smooth(method = "lm", color="black")+
  xlab("rrn copy no. (log2)") +
  ylab("Max. abundance ") +
  ggtitle("Best (HM/LM)") +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  geom_point(aes(color = Phylum),alpha = 0.7, size =3) + 
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

best2

#Foldchnages_HM
best.lm = lm(Fold_changes_best~ log2rrn , data=Best_test)
summary(best.lm) 

#drawing graph in ggplot
Phylum = Best_test$Phylum
best3=ggplot(Best_test, aes(x=log2rrn, y=Fold_changes_best)) +
  geom_smooth(method = "lm", color="black")+
  xlab("rrn copy no. (log2)") +
  ylab("Fold change (abundance)") +
  ggtitle("Best (HM/LM)") +
  geom_point(aes(color = Phylum),alpha = 0.7, size =3) + 
  theme_bw() +
  scale_color_manual(values = c("#0AFD37", "red", "#ffae19",
                                "#4daf4a", "#1919ff","#990000" , "magenta",
                                "#3295ff", "#660066", 
                                "#32ffe7", "black"))+
  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

best3
p1<- multiplot(HM1,HM2,HM3,LM1,LM2,LM3,best1,best2,best3, cols=3)
```

#### Line graphs (Figure 3, Figure 5, Figure 6) 
```{r}
#### Figure 3 (DNA)
Fs3 <- read.csv("/file path/Figure_3.csv",fill = TRUE, header = TRUE, sep = ",")
library(Rmisc)
tgc1 <- summarySE(Fs3, measurevar="Abundance", groupvars=c("Treatment","Day","Phylum", "Treatment1", "Treatment2"))
tgc1

require(grid)
require(plyr)
library(ggplot2)

pd <- position_dodge(0.001) 
a=ggplot(tgc1, aes(x=Day, y=Abundance, group=Treatment)) + 
  
  facet_wrap(~Phylum, ncol = 4, scales = "free_y") +
  
  geom_errorbar(aes(ymin=Abundance-se, ymax=Abundance+se), colour="black", position=pd) +
  geom_line(aes(linetype=Treatment1, colour=Treatment2),position=pd) +
  geom_point(aes(shape=Treatment2,colour=Treatment2))+
  xlab("Day") +
  ylab("Absolute abundance") +
  expand_limits(y=0) +                       
  scale_y_continuous() +        
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.text.x = element_text(size=8, angle=0))  
fig3= a+scale_color_manual(values=c("#4B0082", "#87CEEB")) 
fig3

#Figure 5
Fs5 <- read.csv("/file path/Figure_5.csv",fill = TRUE, header = TRUE, sep = ",")

tgc5 <- summarySE(Fs5, measurevar="Abundance", groupvars=c("Treatment","Day","order_OTU", "Treatment1", "Treatment2"))

library(grid)
library(plyr)
library(ggplot2)

pd <- position_dodge(0.001) 

a=ggplot(tgc5, aes(x=Day, y=Abundance, group=Treatment)) + 
  
  facet_wrap(~order_OTU, ncol = 4, scales = "free_y") +
  
  geom_errorbar(aes(ymin=Abundance-se, ymax=Abundance+se), colour="black", position=pd) +
  geom_line(aes(linetype=Treatment1, colour=Treatment2),position=pd) +
  geom_point(aes(shape=Treatment2,colour=Treatment2))+
  xlab("Day") +
  scale_x_continuous(breaks = c(0,7,14,21,35,50,63))+
  
  ylab("Relative abundance") +
  expand_limits(y=0) +                 
  scale_y_continuous() +        
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.text.x = element_text(size=7, angle=0),
        axis.text=element_text(size=8))  
fig5= a+scale_color_manual(values=c("#4B0082", "#87CEEB")) 
fig5


#Figure 6
Fs6 <- read.csv("file path/Figure_6.csv",fill = TRUE, header = TRUE, sep = ",")
tgc6 <- summarySE(Fs6, measurevar="Abundance", groupvars=c("Treatment","Day","order_OTU", "Treatment1", "Treatment2"))

library(Rmisc)
require(grid)
require(plyr)
require(ggplot2)

pd <- position_dodge(0.001) 
pp=ggplot(tgc6, aes(x=Day, y=Abundance, group=Treatment)) + 

  facet_wrap(~order_OTU, ncol = 4, scales = "free_y") +
  
  geom_errorbar(aes(ymin=Abundance-se, ymax=Abundance+se), colour="black", position=pd) +
  geom_line(aes(linetype=Treatment1, colour=Treatment2),position=pd) +
  geom_point(aes(shape=Treatment2,colour=Treatment2))+
  xlab("Day") +
  ylab("Relative abundance") +
  expand_limits(y=0) +                        
  scale_y_continuous() +        
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        strip.text.x = element_text(size=7, angle=0))  
fig6= pp+scale_color_manual(values=c("#4B0082", "#87CEEB")) 
fig6
```
